% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py (prose drafted 2026-06-25).

% ── 4.1  Controlled-Comparison Design ────────────────────────────────────
% PURPOSE: Establish the single-factor experimental logic that makes H1-H5 valid.
% Model is deepseek-v4-flash (D8) -- NOT MiniMax M3 (stale D6 in scaffold beats).
% ─────────────────────────────────────────────────────────────────────────

\section{Controlled-Comparison Design}

Rather than co-vary model, retrieval, dataset, and storage at once --- as most prior memory
systems do --- this study varies \emph{exactly one} factor, the storage and forgetting
policy, and holds everything else fixed. The logic is that of a single-factor experiment:
if six conditions differ in one and only one component, any between-condition difference is
attributable to that component. Everything that is not the forgetting policy is therefore an
invariant, frozen and shared across all six conditions and all three seeds.

The held-constant factors are enumerated as design invariants. The task stream is the eight
official SWE-Bench-CL sequences, run in their fixed order with no reordering (Invariant~\#1).
The model is frozen: a single \texttt{deepseek-v4-flash} instance serves every model role
across all conditions (the deviation chain from the pre-registered model is disclosed in
Section~\ref{sec:deviations}). Retrieval is the central control --- pure cosine similarity,
identical across all six policies, with no per-policy bonuses or penalties (Invariant~\#5)
--- at a fixed depth $k=5$, same-repository only (Invariant~\#16), with the best-scoring
record injected last to mitigate the lost-in-the-middle effect (Invariant~\#6). The agent
runs at most 20 steps per task (Invariant~\#3), and the embedding payload is the compact
\texttt{[Issue + Final Error + Final Diff]} under the embedder's token cap (Invariant~\#4).
A memory budget cap of ten records (amendment A1) binds the four interior policies. The
single manipulated factor is what the \texttt{MemoryStore} keeps versus forgets between
tasks.

This design buys a specific, bounded kind of validity. Because the model, retrieval,
dataset, and agent loop are constants, the between-policy contrasts (H1--H5) are valid as
statements about forgetting. They are \emph{not}, however, statements about absolute
capability: holding a substituted model and host fixed (Section~\ref{sec:deviations}) means
the absolute resolution rates are not comparable to public SWE-Bench leaderboards. That is a
consequence of the design to be disclosed, not defended --- the comparison is internal by
construction, and its strength is precisely that it does not depend on the absolute numbers.
Figure~\ref{fig:architecture} shows the pipeline with the policy box as the only swappable
component.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
  font=\small,
  box/.style={draw, rounded corners=2pt, align=center, minimum height=9mm,
              minimum width=22mm, inner sep=4pt, thick},
  fixed/.style={box, draw=cFixed},
  agent/.style={box, draw=cAgent, line width=0.6pt},
  flow/.style={-{Stealth[length=2.2mm]}, thick},
  hold/.style={-{Stealth[length=2.2mm]}, thick, dashed, draw=cPolicy},
  node distance=10mm and 16mm,
]
  \node[box]                   (data)   {SWE-Bench-CL\\\scriptsize task sequence};
  \node[agent, right=of data]  (agent)  {Coding Agent\\\scriptsize LangGraph ReAct, $\le$20 steps};
  \node[fixed, right=of agent] (retr)   {Cosine Retriever\\\scriptsize fixed top-$k$, identical};
  \node[fixed, right=of retr]  (store)  {MemoryStore\\\scriptsize FAISS index};
  \node[box, above=of agent]   (llm)    {deepseek-v4-flash\\\scriptsize frozen, all roles};
  \node[box, below=of agent]   (eval)   {eval\_v3 harness\\\scriptsize scores the patch};
  \node[fixed, draw=cPolicy, line width=1pt, above=of store] (policy)
        {Forgetting Policy\\\scriptsize 1 of 6 -- swappable};
  \node[fixed, below=of store] (refl)   {Reflection + Classifier\\\scriptsize 5-type taxonomy};
  \draw[flow] (data)  -- (agent);
  \draw[flow] (agent) -- node[above,font=\scriptsize]{retrieve} (retr);
  \draw[flow] (retr)  -- node[above,font=\scriptsize]{top-$k$} (store);
  \draw[flow] (agent) -- node[right,font=\scriptsize]{LLM calls} (llm);
  \draw[flow] (agent) -- node[right,font=\scriptsize]{patch} (eval);
  \draw[flow] (refl)  -- node[right,font=\scriptsize]{store record} (store);
  \draw[hold] (policy) -- node[right,font=\scriptsize]{maintain / forget} (store);
  \begin{scope}[on background layer]
    \node[draw=cFixed!60, dashed, rounded corners, inner sep=6mm,
          label={[font=\scriptsize, text=cFixed]above:Memory subsystem (retrieval held constant)},
          fit=(retr)(store)(policy)(refl)] {};
  \end{scope}
\end{tikzpicture}
\caption{The memory-equipped coding agent. The agent interacts with a \texttt{MemoryStore}
through a fixed cosine retriever (identical across all six conditions) over the task
sequence; the storage/forgetting \emph{policy} is the single swappable component, one of the
six. The model (\texttt{deepseek-v4-flash}), embedder, retrieval, dataset, and agent loop
are held constant --- the memory subsystem (shaded) operates under retrieval held constant.}
\label{fig:architecture}
\end{figure}
