% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py (prose drafted 2026-06-25).

% ── 4.4  Retrieval as a Fixed Invariant ──────────────────────────────────
% PURPOSE: Retrieval as THE central control -- frozen so differences are storage-only.
% ─────────────────────────────────────────────────────────────────────────

\section{Retrieval as a Fixed Invariant}

The most load-bearing decision in this design is the one that does \emph{not} vary.
Retrieval scoring is pure cosine similarity, identical across all six conditions --- no
bonuses, no penalties, no policy-specific reranking --- and it is enforced in a single
shared code path (\texttt{retriever.py::shared\_retrieve}, Invariant~\#5). Every policy
queries the store the same way; the only thing that differs is what is in the store to be
found. Concretely, a query $q$ and a stored record $m$ are scored by the cosine of their
embeddings:
\begin{equation}
  \label{eq:cosine}
  \operatorname{sim}(q, m) = \frac{\mathbf{e}_q \cdot \mathbf{e}_m}
                                  {\lVert \mathbf{e}_q \rVert \, \lVert \mathbf{e}_m \rVert},
\end{equation}
where $\mathbf{e}_q$ and $\mathbf{e}_m$ are the query and record embeddings; no
policy-specific bonus or penalty ever enters this score.

The retrieval parameters are likewise fixed. Each query returns the top $k=5$ records
(Invariant~\#16), restricted to the same repository, so a sequence never pulls memories from
a foreign project. The retrieved records are injected in relevance order with the
\emph{best} record placed last, nearest the generation point --- the lost-in-the-middle fix,
motivated by the finding that models attend most strongly to the ends of a long
context~\cite{liu2024} (Invariant~\#6).

Freezing retrieval byte-for-byte across conditions is what licenses the causal reading of
the experiment. If two policies retrieve through the same fixed function, then any
difference between them in resolution or footprint cannot be a retrieval difference --- it
can only be a consequence of what each policy chose to store and forget. This is the single
control that turns six otherwise-confounded systems into a clean single-factor comparison
and makes H1--H5 interpretable as statements about forgetting.
