% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py (prose drafted 2026-06-25).
% INTEGRITY: model = deepseek-v4-flash (D8), single model all roles. NOT MiniMax M3.
% Full deviation chain GPT-5.4 -> ... -> deepseek-v4-flash disclosed in Section 6.3.

% ── 4.5  Agent, Benchmark, and Frozen Model ──────────────────────────────
% PURPOSE: Pin the three held-constant factors and disclose the model deviation chain.
% ─────────────────────────────────────────────────────────────────────────

\section{Agent, Benchmark, and Frozen Model}

Three factors are held constant across all 144 runs: the agent, the benchmark, and the
models. The agent is a LangGraph ReAct coding agent that plans, reads and edits files, and
runs tests, with a hard cap of 20 steps per task and a force-fail on overrun
(Invariant~\#3). The benchmark is SWE-Bench-CL~\cite{joshi2025}: all eight official
sequences, in fixed order, with no reordering and no self-generated tasks (Invariant~\#1).
Correctness is scored with the standard \texttt{eval\_v3} harness, so a task counts as
resolved only if the official test suite passes.

The generative model is frozen: a single \texttt{deepseek-v4-flash} instance serves every
model role in the system --- the coding agent, the five-type classifier, the reflection
step, and the CLS consolidation summariser. Using one model for all roles, held identical
across every condition and seed, makes the model a fixed factor: between-policy contrasts
(H1--H5) remain valid even though, as a consequence, the absolute resolution rates are not
comparable to public SWE-Bench leaderboards. This model is the final link in a disclosed
deviation chain from the pre-registered GPT-5.4 (the chain and its rationale are laid out in
Section~\ref{sec:deviations}); \texttt{deepseek-v4-flash} is a light reasoning model, which
has a uniform, between-condition-neutral effect on the token-cost axis that is disclosed
there. Embeddings are computed by a separate local \texttt{nomic-embed-text-v2-moe} model
(768-dimensional), also held constant across conditions.

A structural property of the benchmark bounds how much any of this can matter. Within each
sequence only a minority of tasks share gold-patch files with a predecessor --- the only tasks
for which a stored record could be structurally relevant --- so for most tasks there is, by
construction, little to remember. The per-sequence fractions
(Figure~\ref{fig:interdependence}) range from $0.129$ (scikit-learn) to $0.571$
(pydata/xarray), averaging roughly $0.3$ across the eight sequences: even in the most
interdependent sequence a substantial share of tasks stand alone, and in the least
interdependent sequence the overwhelming majority do. This interdependence ceiling frames the
entire memory question and is revisited in Chapter~\ref{ch:results}.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}
\begin{axis}[width=0.85\textwidth,height=6cm,ybar,bar width=10pt,
  ylabel={Frac.\ tasks with a prior dependency},
  symbolic x coords={sklearn,astropy,sympy,django,pytest,mpl,sphinx,xarray},
  xtick=data,x tick label style={rotate=40,anchor=east,font=\scriptsize},
  ymin=0,ymax=0.62,nodes near coords,every node near coord/.append style={font=\tiny}]
\addplot coordinates {(sklearn,0.129)(astropy,0.143)(sympy,0.163)(django,0.286)(pytest,0.389)(mpl,0.394)(sphinx,0.512)(xarray,0.571)};
\end{axis}
\end{tikzpicture}
\caption{Structural task interdependence per sequence: the fraction of tasks whose gold-patch
files overlap an earlier task. Only a minority depend on a predecessor (mean
\textasciitilde 0.3), a property of the benchmark.}
\label{fig:interdependence}
\end{figure}
