% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py (prose drafted 2026-06-25).
% INTEGRITY: CL-F1 reported as resolve-rate PROXY (anchor-probe demoted, A5). TOST/SESOI was
% PRE-REGISTERED but NOT computed -- non-inferiority reported via bounded CIs + power, not a
% TOST verdict (advisor landmine #3). Cite only bib keys (no Lakens-TOSTER).

% ── 4.6  Metrics and Pre-Registered Statistical Plan ─────────────────────
% PURPOSE: Define outcome/cost metrics and the effect-size-first, pre-registered plan.
% ─────────────────────────────────────────────────────────────────────────

\section{Metrics and Pre-Registered Statistical Plan}
\label{sec:metrics}

\paragraph{Outcome metric.}
The primary outcome is CL-F1, the continual-learning resolution measure that the
pre-registration defined as a combination of \emph{plasticity} (solving the current task)
and \emph{stability} (retaining competence on earlier tasks), the latter measured by
anchor-probe re-evaluation. In this thesis CL-F1 is reported as a \emph{resolution-rate
proxy}: the full anchor-probe instrumentation was not run at matrix scale, so the stability
dimension is addressed separately through a targeted construct-validity sub-study
(amendment A5; Section~\ref{sec:h5}) rather than folded into the headline metric. Readers
should read CL-F1 here as the per-condition resolution rate, with the stability question
treated explicitly and its limitation disclosed in Threats to Validity
(Section~\ref{sec:threats}).

\paragraph{Cost metrics.}
Cost is reported on a two-axis Pareto frontier (Table~\ref{tab:metrics}): a \emph{compute}
axis (total tokens per run, a provider-independent proxy under deviation D3) and a
\emph{footprint} axis (memory tokens carried per task, plus retrieval-index growth). The
footprint axis is the A4 reframing of the original H1b efficiency hypothesis: because
retrieval depth is fixed, storage pruning cannot reduce the prompt budget, so the honest
place to look for an efficiency win is footprint, not total-token savings.

\paragraph{Inferential plan.}
The analysis is pre-registered and effect-size-first, reporting estimation --- intervals and
effect sizes --- ahead of dichotomous significance, in the spirit
of~\cite{cumming2014} and~\cite{wasserstein2019}. The primary confirmatory test is the
Wilcoxon signed-rank test on $N=8$ sequence-level means, comparing each policy against Full
Memory with Holm correction over the five pre-registered contrasts (Invariant~\#11); the
sequence, not the task, is the unit of analysis, to avoid pseudo-replication. Effect size is
the rank-biserial correlation $r_{rb}$ (Invariant~\#12, deliberately not Cohen's $d$ or
Cliff's $\delta$), and confidence intervals are BCa bootstrap with 5000 resamples
(Invariant~\#15). Non-inferiority was pre-registered against a smallest-effect-size-of-
interest of $\pm0.03$ CL-F1; the thesis reports it through the bounded paired-difference
interval and the design's detectable-effect (power) profile rather than as a separate
equivalence-test verdict, and does not switch tests to manufacture significance. At the task
level, a generalised linear mixed model (binomial/logit) with a crossed random effect on
\texttt{task\_id} (Invariant~\#14) complements the sequence-level test, and the
helpful-versus-harmful feature analysis uses PR-AUC with a VIF collinearity check rather
than accuracy or ROC-AUC, given the low positive-class rate (Invariant~\#13).

\begin{table}[htbp]
  \centering
  \caption{Metrics and their roles. Numeric values are reported in
  Chapter~\ref{ch:results} from the trustworthy rerun.}
  \label{tab:metrics}
  \begin{tabular}{lll}
    \toprule
    Metric & Definition & Role / axis \\
    \midrule
    CL-F1 (proxy)   & per-condition resolution rate         & primary outcome \\
    Footprint       & memory tokens carried per task         & efficiency (footprint axis) \\
    Total tokens    & tokens per run (D3 proxy)              & efficiency (compute axis) \\
    Tool calls      & agent tool invocations per task        & behavioral (H4) \\
    Memory lift     & full minus no-memory resolution        & interdependence (E7) \\
    \bottomrule
  \end{tabular}
\end{table}
