% Drafted 2026-06-25. NEW methodological contribution (not in the original Typst scaffold).
% Sources: docs/ab-gate-findings-2026-06-24.md, docs/amended-gate-2026-06-24.md,
% results/ab_gate/amended_cceb325_result.json, results/manifest/freeze.json (sha cceb325).
% INTEGRITY: NEVER state "the original gate passed". Original gate STOPped -> root cause ->
% hotfix cceb325 -> amended (pre-registered) gate -> GO. Outcome numbers reported in 5.0.

% ── 4.7  Instrument Validation ───────────────────────────────────────────
% PURPOSE: Document the pre-registered A/B health check of the agent's own tooling,
%          its STOP, the root-caused fix, and the amended re-validation.
% ─────────────────────────────────────────────────────────────────────────

\section{Instrument Validation}
\label{sec:instrument}

A controlled comparison is only as trustworthy as the instrument that produces its data. A
trajectory audit of an early full-matrix run revealed that the agent's own file tools were
defective --- \texttt{read\_file} returned no line ranges and truncated at a few kilobytes,
and \texttt{edit\_file} mishandled path normalisation --- which meant that an apparent
between-policy result could be an artefact of broken tooling rather than of forgetting.
Before trusting any matrix, we therefore ran a pre-registered A/B health check of the
instrument itself, and the result was decisive enough to halt and re-engineer before
re-running. We report the procedure here and its verified outcome in
Section~\ref{sec:instrument-results}.

\paragraph{A/B design.}
The check is a 36-cell paired A/B: two sequences (\texttt{pytest}, \texttt{scikit-learn})
$\times$ three policies (\texttt{no\_memory}, \texttt{full\_memory}, \texttt{recency\_prune})
$\times$ three seeds $\times$ two tool modes (\emph{legacy}, the pre-fix tooling, and
\emph{fixed}, the post-fix tooling). Everything other than the tool mode is held constant,
and every cell is run in both modes so the comparison is fully paired.

\paragraph{The original gate stopped.}
The pre-registered health gate returned \textbf{STOP}. Two criteria failed: a non-zero count
of edit path/index failures, and an edit-failure ratio above the pre-registered $0.15$
threshold. Root-cause analysis attributed the failures to a real bug in the path-handling
fix: \texttt{edit\_file} normalised the diff's paths but then compared them against the raw,
un-normalised \texttt{path} argument, so a perfectly valid edit to an absolute container path
was rejected as a spurious ``security'' violation. The overwhelming majority of the security
rejections --- 77 of 78 --- were this single bug firing on same-file edits (the one remaining
rejection was a genuine cross-file diff, correctly refused). The defect was fixed (commit
\texttt{cceb325}), with regression tests covering the absolute-path and double-slash cases.

\paragraph{A principled amendment, not a moved goalpost.}
After the fix, the residual edit failures are \emph{model-quality} failures --- the frozen
model occasionally emits a malformed unified diff or an incorrect path --- which no
instrument can prevent and which the agent recovers from via retry. We therefore amended the
gate on first principles, \emph{before} collecting any re-run data, and submitted the
amendment for external (Codex) review. The amendment retires the absolute $0.15$
total-ratio criterion --- which conflates instrument health with a fixed property of the
model --- and replaces it with two principled checks: (i) \emph{instrument-attributable}
failures must fall to zero, and (ii) the \emph{model-quality} failure rate in the fixed arm
must not exceed that in the legacy arm (a non-regression guard against the fix silently
degrading diff quality), alongside the unchanged completeness, pairing, provenance, and
token-inflation checks. The instrument-versus-model distinction is defined on first
principles, not fitted to the data: model-quality diff errors are a property of the frozen
model held constant across all conditions and seeds --- exactly the logic by which the
model substitution itself (deviations D1--D8, Section~\ref{sec:deviations}) is
non-confounding for between-policy contrasts. The amended gate's verdict on fresh
post-hotfix data, and the supporting numbers, are reported in
Section~\ref{sec:instrument-results}; the full 144-run matrix was executed only after that
gate returned GO, on the fixed code at the frozen experiment commit \texttt{cceb325}.
