% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py.

% ── 7.1  Conclusion ──────────────────────────────────────────────────────
% PURPOSE: Restate the thesis and recap the metric → trade-off in one tight pass.
% MOVE: Restate + metric → trade-off recap (per the storytelling template).
% BEATS: see ordered bullets below — what we measured, the result, the reframe, the contribution.
% FIGURES/TABLES: none (cross-ref fig:pareto-footprint from Ch5/6).
% CITATIONS: none new (xiong2025 if interdependence is cited).
% SOURCE-DOC: README grounded-facts; docs/defense-qa.md (Q5).
% ─────────────────────────────────────────────────────────────────────────

\section{Conclusion}

This thesis set out to measure, rather than to advocate, what proactive forgetting
does to an AI coding agent that learns across a sequence of tasks. We held a single
frozen model (\texttt{deepseek-v4-flash}) and a single retrieval rule constant --- pure
cosine similarity, identical across all six policies (Invariant~\#5) --- so that the
storage-and-forgetting policy was the only moving part. Six policies, spanning the
No-Memory and Full-Memory brackets together with four pruning and consolidation schemes,
were run across all eight SWE-Bench-CL sequences at three seeds, yielding 144 controlled
runs and 4{,}914 task observations. Performance was read on the CL-F1 metric, which in
this single-frozen-model setting reduces to a resolved-rate proxy; it captures whether the
agent fixed each task, not the shape of a forgetting curve, and we have been deliberate
throughout not to overclaim the latter.

The central result is a \emph{powered null} that survives on the trustworthy data. No
forgetting or consolidation policy beat full-memory accumulation: across the five
pre-registered Wilcoxon contrasts against \texttt{full\_memory} (Invariant~\#11), every
contrast was non-significant after Holm correction, and every $95\%$ bias-corrected
bootstrap confidence interval on the paired difference crossed zero. The same pattern
holds in the task-level binomial GLMM (Invariant~\#14), where no policy coefficient
reaches significance. Whatever advantage carrying more memory confers is small and
bounded: \texttt{full\_memory} resolved $44.5\%$ of tasks against $42.9\%$ for
\texttt{no\_memory}, a difference of roughly $1.6$ percentage points that is statistically
indistinguishable from zero and below the resolution of an $N=8$ sequence-level design ---
which is underpowered for sub-three-point effects. We therefore report a bound, not a
benefit: at this scale, accumulating an agent's full task history buys at most a few points
of resolved rate, and may buy nothing.

Where forgetting visibly wins is the second axis. The pruning policies matched
full-memory's resolved rate while carrying roughly half the per-task memory footprint
(about $600$ tokens against full-memory's $1{,}247$; see
Figure~\ref{fig:pareto-footprint}), and the No-Memory bracket holds the zero-footprint
corner. This advantage is specific: it lives on the memory-footprint axis, not on total
compute. End-to-end token totals and tool-call counts were essentially flat across all six
policies --- the cost of these runs is dominated by the agent's own reasoning loop, which
forgetting does not touch. Forgetting, in other words, is close to free on the dimension it
governs and irrelevant on the dimension it does not; the headline trade-off is non-inferior
performance at a materially lighter memory footprint.

These findings reframe the question. A frozen retrieval rule and a powered null together
suggest that ``how much should an agent remember'' is less an always-on accumulation
default than a problem of targeting. Only a minority of SWE-Bench-CL tasks --- a structural
property of the benchmark --- genuinely depend on a predecessor, and even those that do see
heterogeneous and largely null lift from carrying memory; structural overlap does not
predict where memory helps~\cite{xiong2025}. Compounding this, agent effectiveness declines
with position in a sequence regardless of policy: later tasks are harder for the agent than
earlier ones, and no amount of accumulated memory rescues that trajectory. Not every task
needs memory, and the tasks that might are not reliably the ones that store the most.

The contribution of this work is accordingly a \emph{measurement} and a reusable
\emph{template}, not a new memory policy. By fixing retrieval, bracketing the design
between No-Memory and Full-Memory, separating the footprint and compute axes of cost, and
quantifying task interdependence as a benchmark property, the study isolates the
forgetting-policy effect cleanly enough to detect its absence with confidence --- and
leaves behind a 144-run, fully provenanced artifact on which the next, better-powered
comparison can build.
