% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py.

% ── 7.2  Future Work ─────────────────────────────────────────────────────
% PURPOSE: Each future item traces a NAMED limitation from §6.4 to a concrete next study.
% MOVE: Future-work-per-bottleneck (each item closes a named threat to validity).
% BEATS: see ordered bullets below — cross-model · anchor-probe · decontamination · cross-repo.
% FIGURES/TABLES: none.
% CITATIONS: SWE-rebench 2505.20411 [VERIFY+ADD], SWE-MERA 2507.11059 [VERIFY+ADD].
% SOURCE-DOC: docs/threats-to-validity.md; docs/a5-anchor-probe-decision.md; AMENDMENTS (D-0.4).
% ─────────────────────────────────────────────────────────────────────────

\section{Future Work}

Every limitation catalogued in the threats-to-validity discussion (\S\ref{sec:threats}) maps onto a concrete follow-up study. We frame these as the natural continuations of a deliberately narrow controlled comparison rather than as defects of the present design: each one relaxes a single frozen scope decision and asks whether the powered null reported here---no memory policy beats full accumulation, and performance declines with sequence position regardless of policy---persists under a broader regime.

\paragraph{Cross-model replication.}
The clearest external-validity threat is that the entire matrix runs on a \emph{single} frozen agent model, \texttt{deepseek-v4-flash}. Holding the model constant is exactly what licenses the between-policy comparisons (the model is a fixed factor across all six conditions and three seeds), but it leaves open whether the policy-null and the position decline are properties of memory management in general or artefacts of one model's behaviour. A natural next study replicates the top contrasts---full memory, the aggressive pruners, and no memory---on a small set of structurally different agents (for example a reasoning-tuned model and a larger long-context model), reporting the \emph{direction} of transfer rather than re-running the full inferential design. If the position decline survives a model with a substantially larger context window, it becomes much harder to attribute to one model's context handling.

\paragraph{Cross-repository retrieval.}
The main experiment restricts retrieval to same-repository memories (Invariant~\#16), so the agent never draws on experience from one codebase while solving a task in another. This bounds the external validity of the forgetting policies to the within-project setting. Relaxing the same-repo constraint would let memory accumulate and be retrieved across repositories, changing both the volume of competing items and the structure of interdependence between tasks. Whether content-aware forgetting earns its keep there---where the signal-to-noise ratio of retrieved items is lower and naive accumulation is more likely to dilute the context---is precisely the regime in which a pruning policy could plausibly separate from full memory, and it is the most promising place to look for the effect this study did not find.

\paragraph{The full anchor probe.}
Our primary outcome, CL-F1, is operationalised as a per-task resolve-rate proxy, not as a forgetting curve over a fixed anchor set. The stability instrument therefore cannot, on its own, distinguish a policy that genuinely preserves earlier solutions from one that simply re-derives them. The A5 anchor-probe sub-study (\S\ref{sec:threats}) was a deliberately small discrimination test confirming that the construct \emph{can} be exercised, run on a hand-selected, selection-maximised set of cells and therefore reported without a $p$-value. The follow-up is the full probe: a pre-registered anchor set drawn from interdependent tasks across the high-overlap sequences, evaluated for each policy at several seeds, so that retention is measured directly and tested inferentially rather than read off a resolve-rate proxy. This would convert the present construct-validity caveat into a measured quantity.

\paragraph{Tier-2 gold labels for the feature analysis.}
The helpful/harmful classification of retrieved memories (Invariant~\#13) passed its collinearity screen---semantic similarity, retrieval rank, and memory age all show acceptable variance-inflation factors---but the PR-AUC model itself remains deferred. The Tier-1 weak-labelling heuristic produced an all-neutral target with no positive class, so the classifier cannot yet be fit. Completing this analysis requires Tier-2 manual gold labels: human adjudication, on a sampled subset of the task--memory pairs, of whether an injected memory actually contributed to or harmed the solution. With a real positive class the pre-registered PR-AUC analysis becomes estimable, and the question of \emph{which} memory features predict usefulness---rather than merely which are non-collinear---can finally be answered.

\paragraph{The position decline as the larger lever.}
The most consequential finding for future work is not about any storage policy at all. In the task-level GLMM, the storage policy coefficients are indistinguishable from zero, while sequence position carries a large, significant negative effect: the agent gets worse as a sequence progresses, and accumulating memory does not rescue it. This points away from \emph{what to remember} and toward \emph{how the working context degrades}---the context-dilution and lost-in-the-middle dynamics that storage-side pruning cannot touch \citep{liu2024}. The higher-leverage research programme is therefore on the retrieval and context-construction side: how items are ranked, how many are injected, and how the prompt is assembled as a sequence lengthens. The contribution of this thesis is partly to redirect attention there: on this benchmark and this model, the forgetting policy is a footprint lever, while the position decline is where the performance is won or lost.
