% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py (prose drafted 2026-06-25).
% ── 4.2.6  CLS Consolidation ─────────────────────────────────────────────

\subsection{CLS Consolidation}
\label{sec:cls}

CLS Consolidation forgets by \emph{compression} rather than deletion. On a fixed schedule
--- every $k=5$ tasks (Invariant~\#9) --- it clusters the older memories with DBSCAN, forms
a cluster only where at least three records lie within $0.70$ cosine similarity of one
another, and replaces each qualifying cluster with a single LLM-written summary. This models
the slow neocortical consolidation of Complementary Learning Systems
theory~\cite{mcclelland1995}: stable, repeated structure is abstracted into a compact gist
while one-off details fade. When no cluster qualifies, the policy falls back to Type-Aware
Decay (Section~\ref{sec:type-decay}) at the cap of ten (A1), so it never simply overflows.

The expectation, from the mechanism, is that on SWE-Bench-CL this policy is largely
\emph{inert}: the same-repository memories within a sequence are diverse --- each a different
issue --- so a cluster of three or more within the $0.70$ cosine threshold rarely forms and
consolidation seldom fires. This is a gate-3 \emph{indication}, not yet a matrix-wide finding:
in the gate-3 validation the policy produced zero consolidation events on both probed
sequences (amendment A3), reducing empirically to its Type-Aware Decay fallback. Whether this
holds across the full matrix --- the matrix-wide consolidation count --- is reported with the
H3 results (Section~\ref{sec:h3}) \texttt{[RESULT: pending runs\_144\_seq\_cceb325]}. We
deliberately did \emph{not} relax the $0.70$ threshold to force clustering: that parameter is
frozen (Invariant~\#23), and lowering it to manufacture a positive result would merge
dissimilar memories into incoherent summaries. If the matrix-wide count confirms the gate-3
indication, H3 is reported as a negative, not-applicable result --- itself a finding about how
compressive forgetting behaves on genuinely heterogeneous coding experience.
