% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py (prose drafted 2026-06-25).
% ── 4.2.5  Type-Aware Decay ──────────────────────────────────────────────

\subsection{Type-Aware Decay}
\label{sec:type-decay}

Type-Aware Decay is the content-aware policy at the top of the ladder's interior. Each
record carries a retention score that combines how old it is, how fast records of its kind
should decay, and how often it has been retrieved; when the store is over the cap of ten
(A1), the lowest-scoring records are evicted. The score is the multiplicative
Anderson--Schooler form~\cite{anderson1991} (Invariant~\#8):
\[
  \mathrm{score} = \mathrm{base}\times\mathrm{age}^{-d(\mathrm{type})}
  \times(1+\mathrm{retrieval\_count})^{0.5}.
\]
Age pulls the score down, retrieval pulls it back up (the square root giving diminishing
returns), and the per-type exponent $d(\mathrm{type})$ sets how quickly each kind of record
ages. The exponent is keyed to the five-type taxonomy --- \texttt{architectural},
\texttt{api\_change}, \texttt{bug\_fix}, \texttt{test\_update}, \texttt{config}
(Invariant~\#7) --- so durable structural knowledge decays slowly while ephemeral records
fade fast. Forgetting is thus driven by content and use, not by age alone.

Crucially, $d(\mathrm{type})$ is \emph{frozen at design time} from the theoretical values
and is never calibrated or re-tuned per type from pilot data. This is a pre-registration
integrity choice: fitting one decay exponent per type on a handful of sequences would
overfit and turn a hypothesis test into a curve fit. The policy as evaluated therefore
tests the cognitive hypothesis, not a benchmark-tuned variant of it.
