% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py (prose drafted 2026-06-25).
% NOTE: Shen (2602.21611) and EXPEREPAIR (2506.10484) were [VERIFY+ADD] in the scaffold but
% are NOT in references.bib -- the coding-agent-memory pattern is stated generically here
% rather than with uncited specifics. Add to bib + cite if these works are to be named.

% ── 3.2  Continual Learning Benchmarks and Coding-Agent Memory ───────────
% PURPOSE: Survey CL/memory benchmarks and coding-agent memory systems thematically.
% ─────────────────────────────────────────────────────────────────────────

\section{Continual Learning Benchmarks and Coding-Agent Memory}

\subsection{Continual-learning and memory benchmarks}

The benchmark this thesis builds on, SWE-Bench-CL~\cite{joshi2025}, arranges real GitHub
issues into ordered, same-repository sequences with genuine dependency structure --- the
substrate a continual-learning study of coding agents needs. What it supplies is the task
stream; what it does not supply is a protocol for comparing \emph{forgetting policies},
which this thesis adds on top of it. Broader memory-evaluation suites sit further from the
target: MemBench~\cite{membench2025} assembles a comprehensive battery for agent memory but
is oriented to conversational and question-answering interaction rather than
coding-with-forgetting, and Evo-Memory~\cite{wei2025evomemory} benchmarks self-evolving
memory across tasks but is QA-centric and does not isolate the storage policy under fixed
retrieval. The common gap is that no existing benchmark pairs coding-agent forgetting with
retrieval held constant; we adapt SWE-Bench-CL to exactly that protocol.

\subsection{Coding-agent memory systems}

A parallel line equips coding and program-repair agents with episodic or semantic memory of
past fixes and reports that it improves resolution. Two features recur across these systems
and matter for how their results should be read. First, they are framed as
\emph{superiority} claims --- memory helps --- rather than as the non-inferiority question
of whether a lighter store can match a heavier one. Second, they typically vary the
retrieval mechanism (its granularity, its ranking, or what unit is indexed) alongside the
memory contents, so a reported gain cannot be attributed to storage as opposed to search.
Both features motivate the design here: fix retrieval, vary only the storage and forgetting
policy, and ask the non-inferiority question --- can an agent forget aggressively and lose
nothing on task resolution --- rather than assuming more memory is better.
