% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py.

% ── Appendix A  Reproducibility and Provenance ───────────────────────────
% PURPOSE: Give a reader everything needed to reproduce the matrix and audit provenance.
% MOVE: none (reference appendix — structured disclosure, not narrative).
% BEATS: see ordered bullets below — pre-reg, archive, seeds, environment, model, logging, reversibility.
% FIGURES/TABLES: cross-ref tab:deviations (§6.3).
% CITATIONS: none new.
% SOURCE-DOC: AMENDMENTS.md (D2/D5/D6); CLAUDE.md (provider config, logging schemas v5 §11).
% ─────────────────────────────────────────────────────────────────────────

\section{Reproducibility and Provenance}
\label{sec:reproducibility}

This appendix lists what a replicator needs to re-derive the matrix and audit its provenance.
Each item names the concrete artifact to consult. The reported matrix is
\texttt{runs\_144\_seq\_cceb325} (data) analysed into \texttt{results\_seq\_cceb325}
(tables, GLMM, plots); the full deep audit returned $0$ flags.

\begin{description}
  \item[Pre-registration.] The frozen design is \texttt{THESIS\_FINAL\_v5} ($26$ frozen
    decisions, $16$ invariants). Every change from it is logged as a dated deviation or
    amendment (D1--D8, A1--A7) in \texttt{AMENDMENTS.md} and summarised in
    Table~\ref{tab:deviations} (Section~\ref{sec:deviations}).
  \item[Frozen experiment commit.] All $144$ cells were produced by the agent code frozen at
    commit \texttt{cceb325} (\texttt{git diff cceb325..HEAD -- src/agents/} is empty), so the
    science is identical across the matrix. The instrument hotfix that defines \texttt{cceb325}
    is described in Section~\ref{sec:instrument}; orchestration changes made during execution
    (per-task Docker image cleanup, sequence-phased scheduling) are disk-management only and do
    not alter what is computed.
  \item[Design and seeds.] Six policies $\times$ eight official SWE-Bench-CL sequences $\times$
    three seeds $=144$ runs (A7-restored). Seeds are indexed $1$--$3$ and applied identically
    across all policies and sequences; the cell directory names encode policy, sequence, and
    seed (e.g.\ \texttt{cls\_consolidation\_django\_seed2}). The three per-seed CL-F1 values per
    cell are retained in \texttt{aggregated/sequence\_aggregates.json}.
  \item[Model provenance.] The executed model is a single frozen
    \texttt{deepseek-v4-flash} on OpenCode serving \emph{all} roles --- agent, $5$-type
    classifier, reflection, and CLS summariser (D8). It is the endpoint of a disclosed chain of
    OpenAI-compatible substitutions away from the pre-registered GPT-5.4 (Section~\ref{sec:deviations});
    the intermediate MiniMax~M3 run (D6) was discarded with no usable data. Because the model is
    held constant across all conditions, it is a fixed factor; absolute resolution rates are not
    comparable to GPT-5.4 or public leaderboards. A rotating key pool fails closed on quota
    exhaustion so balance depletion is never recorded as a false \texttt{resolved=0}.
  \item[Computational environment.] An x86\_64 cloud fleet with prebuilt SWE-bench instance
    images (D5); embeddings from a local Ollama \texttt{nomic-embed-text-v2-moe} ($768$-d,
    deterministic) (D2); cost recorded as a token-count proxy because the provider is flat-rate
    (D3). Reasoning tokens are stripped deterministically before parsing, embedding, and
    re-injection, so the memory payload is reasoning-free (D6a/D8a).
  \item[Per-run logging.] Each run records its configuration and model identity alongside
    \texttt{task\_results.jsonl}, \texttt{memory\_events.jsonl}, per-task trajectories, and
    before/after memory snapshots (v5~\S11 schemas). These are the inputs the audit and the
    analysis pipeline consume.
  \item[Re-derivation.] The full analysis is reproduced with
    \texttt{python -m scripts.run\_analysis --stage all --runs-dir runs\_144\_seq\_cceb325
    --out results\_seq\_cceb325}; the task-level GLMM is a separate stage
    (\texttt{src.analysis.glmm.run\_task\_level\_analysis}).
  \item[Reversibility.] The provider is selected entirely through \texttt{.env}; chat and
    embedding clients are separate (no global \texttt{OPENAI\_BASE\_URL}), so the model and
    embedder deviations are configuration-only and reversible to GPT-5.4/OpenAI without code
    changes.
\end{description}
