% ── 5.2  H1: Resolution and the Performance--Cost / Footprint Frontiers ──────
% PURPOSE: The signature analysis -- whether forgetting matches full memory on the
%          CL-F1 proxy, and how the two cost axes (compute vs footprint) behave.
% Primary test = Wilcoxon (Invariant #11); TOST/SESOI was NOT computed -- non-inferiority
% is read from bounded CIs, not a TOST verdict. Data = runs_144_seq_cceb325 (deepseek-v4-flash).
% ─────────────────────────────────────────────────────────────────────────

\section{H1: Resolution and the Performance--Cost / Footprint Frontiers}
\label{sec:h1}

This section reports the thesis's primary contrast: whether any forgetting policy differs
from full-memory accumulation on the resolution proxy, and how cost behaves on the compute
and footprint axes. Because retrieval is held constant across all six conditions, the
contrast is attributable to \emph{what} each policy keeps rather than to \emph{how} memory
is searched. The headline is a \emph{powered null}: across the trustworthy
$144$-run matrix (\texttt{runs\_144\_seq\_cceb325}: $4{,}914$ task rows, $2{,}175$
resolved, $44.3\%$ overall), no forgetting policy differs significantly from
\texttt{full\_memory} on the resolution proxy, the memory benefit over carrying no memory
at all is small and statistically indistinguishable from zero, and aggressive forgetting is
\emph{footprint-free} --- it matches full accumulation at roughly half the carried memory
without costing compute. Throughout, the value axis is CL-F1 measured as a
\emph{resolved-rate proxy}: it captures end-of-sequence task success, not a fine-grained
forgetting curve, a limitation we revisit in the construct-validity discussion
(Section~\ref{sec:threats}) and probe directly in the A5 anchor sub-study.

\subsection{The primary test}

The pre-registered primary analysis is a Wilcoxon signed-rank test on $N=8$
sequence-level means, comparing each of the five other policies against
\texttt{full\_memory}, with Holm correction across the five contrasts
(Section~\ref{sec:metrics}). Effect size is rank-biserial $r_{rb}$ and the interval is the
95\% BCa bootstrap CI on the paired difference in resolution rate.
Table~\ref{tab:main-results} reports the five contrasts. \emph{No contrast is
significant}: every Holm-corrected $p$ equals $1.000$ and every BCa interval on the paired
difference crosses zero. The point estimates are tightly clustered --- the six policy means
span only $0.427$ (\texttt{type\_aware\_decay}) to $0.449$ (\texttt{cls\_consolidation}),
with \texttt{full\_memory} at $0.445$ sitting in the middle of the pack rather than at the
top. There is thus no evidence that any forgetting policy underperforms full accumulation
on the resolution proxy, and equally no evidence that any beats it.

\begin{table}[htbp]
  \centering
  \caption{H1 primary test. Wilcoxon signed-rank on $N=8$ sequence means, each policy vs
  \texttt{full\_memory}, Holm-corrected over the five contrasts. Effect size is
  rank-biserial $r_{rb}$; the interval is the 95\% BCa bootstrap CI on the paired
  difference in resolution rate. All five contrasts are non-significant after Holm
  correction, and every CI crosses zero (\texttt{runs\_144\_seq\_cceb325}).}
  \label{tab:main-results}
  \begin{tabular}{lrrrrlc}
    \toprule
    Contrast (vs \texttt{full\_memory}) & $W$ & $p$ & Holm $p$ & $r_{rb}$ & 95\% BCa CI & Sig. \\
    \midrule
    random\_prune      & 17.00 & 0.945 & 1.000 & $-0.056$ & $[-0.042,\ 0.027]$ & No \\
    recency\_prune     & 16.00 & 0.844 & 1.000 & $-0.111$ & $[-0.053,\ 0.038]$ & No \\
    type\_aware\_decay & 10.00 & 0.297 & 1.000 & $-0.444$ & $[-0.063,\ 0.016]$ & No \\
    cls\_consolidation & 13.00 & 0.938 & 1.000 & $-0.071$ & $[-0.042,\ 0.034]$ & No \\
    no\_memory         & 10.00 & 0.578 & 1.000 & $-0.286$ & $[-0.053,\ 0.023]$ & No \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Bounding and powering the result}

A non-significant result is informative only if its magnitude is bounded. We therefore
report, alongside the primary test, the aggregate \texttt{full\_memory}-minus-\texttt{no\_memory}
difference with its bootstrap interval, in the estimation spirit of \cite{cumming2014} and
\cite{wasserstein2019} --- as an interval that bounds what the null does or does not rule
out, \emph{not} as a second significance verdict, and with no test-switching to manufacture
significance (Invariant \#11).

Carrying full memory raised the resolution proxy from $0.429$ (\texttt{no\_memory}) to
$0.445$ (\texttt{full\_memory}), a difference of $+1.6$ percentage points on the
across-sequence means (paired median $\Delta$ of $+3.4$ percentage points). This benefit is
not statistically distinguishable from zero: the contrast is non-significant after Holm
correction (Holm $p = 1.000$, $r_{rb} = -0.286$) and its BCa bootstrap interval on the
paired difference crosses zero. The benefit, in other words, is small and positive but
unconfirmed. We read this as an \emph{upper bound} rather than a power-calibrated null: the
$N=8$ sequence-level design does not have the resolution to confirm effects of this size,
so a sub-three-percentage-point memory benefit can be neither asserted nor ruled out. What
the data \emph{can} say is that memory does not buy a large advantage here, and that no
forgetting policy gives up the small advantage it does buy.

\subsection{The two-axis frontier: compute versus footprint}

The efficiency contribution (H1b) is read on two separate axes, because forgetting can
behave very differently on each.

\paragraph{Compute.}
On total token cost (Figure~\ref{fig:pareto-compute}) the six policies barely separate ---
a near-null. Total tokens per run rise only from $5.74$M for \texttt{no\_memory} to $6.09$M
for \texttt{full\_memory}, and the five memory-carrying policies fall within roughly $2\%$
of one another ($5.96$--$6.09$M); the full spread across all six is about $6\%$. This is the
expected consequence of an agent-loop-dominated cost: each task runs up to $20$ ReAct steps
whose growing conversation is resent every step (a near-constant $\approx 21$ tool calls per
task across all policies; Section~\ref{sec:h4}), so the memory payload is a small fraction of
the per-task token bill. The practical implication is blunt: \emph{pruning does not reduce
total compute}. Whatever the case for forgetting, it is not made on the compute axis.

\paragraph{Footprint.}
The memory-footprint axis (Figure~\ref{fig:pareto-footprint}) measures the memory tokens
carried per task --- the quantity that governs storage and retrieval-index growth over a
long-running agent's lifetime, and the axis on which the pruning policies are designed to
differ from \texttt{full\_memory}. Here the separation is real. \texttt{full\_memory}
carries $1{,}247$ tokens per task; the pruning policies carry roughly half of that
($598$ for \texttt{type\_aware\_decay}, $601$ for \texttt{random\_prune}, $604$ for
\texttt{recency\_prune}, $704$ for \texttt{cls\_consolidation}), and \texttt{no\_memory}
holds the zero-footprint corner. Because all of these sit at statistically
indistinguishable resolution rates (Table~\ref{tab:main-results}), the pruners are not
Pareto-dominated by \texttt{full\_memory}: they match its task success at roughly half the
carried footprint. Forgetting, on this benchmark, is \emph{footprint-free}.

\begin{figure}[htbp]\centering
\begin{tikzpicture}\begin{axis}[width=0.78\textwidth,height=6cm,xlabel={Total tokens per run (M)},
  ylabel={CL-F1 (resolve rate)},only marks,xmin=5.6,xmax=6.25,ymin=0.42,ymax=0.455,
  mark size=2.2pt]
% all six policy means; only the two extremes are labelled -- the four memory-carrying
% policies form one tight cluster (the figure's point: a near-null on compute).
\addplot[only marks,mark=*,color=cPolicy] coordinates {(5.74,0.429)};
\node[font=\scriptsize,anchor=east] at (axis cs:5.74,0.429) {no\_memory};
\addplot[only marks,mark=*,color=cFixed] coordinates {(5.96,0.427) (6.05,0.437) (6.06,0.447) (6.08,0.449)};
\addplot[only marks,mark=*,color=cAgent] coordinates {(6.09,0.445)};
\node[font=\scriptsize,anchor=west] at (axis cs:6.09,0.445) {full\_memory};
\end{axis}\end{tikzpicture}
\caption{Performance versus compute (total tokens per run). The six policy means overlap
within roughly $6\%$ on the compute axis (the five memory-carrying policies within $\approx
2\%$) --- a near-null: pruning does not reduce total compute, because cost is dominated by the
agent loop rather than the memory payload. Only the two extremes are labelled; the clustered
points are, left to right, \texttt{type\_aware\_decay}, \texttt{recency\_prune},
\texttt{random\_prune}, \texttt{cls\_consolidation}.}
\label{fig:pareto-compute}\end{figure}

\begin{figure}[htbp]\centering
\begin{tikzpicture}\begin{axis}[width=0.78\textwidth,height=6cm,xlabel={Memory footprint (tokens/task)},
  ylabel={CL-F1 (resolve rate)},only marks,xmin=-150,xmax=1400,ymin=0.42,ymax=0.455,
  nodes near coords,point meta=explicit symbolic,every node near coord/.append style={font=\scriptsize,anchor=south}]
\addplot+[mark=*] coordinates {(0,0.429)[no\_mem] (598,0.427)[type] (601,0.447)[random] (604,0.437)[recency] (704,0.449)[cls] (1247,0.445)[full]};
\end{axis}\end{tikzpicture}
\caption{Performance versus memory footprint. \texttt{full\_memory} (far right, 1{,}247 tok) is matched on
CL-F1 by the pruners at roughly half the footprint (\textasciitilde 600 tok); \texttt{no\_memory} holds the zero-footprint corner.}
\label{fig:pareto-footprint}\end{figure}

\paragraph{The combined claim.}
Read together, the value axis and the two cost axes give a single, defensible statement:
aggressive forgetting matches full-memory accumulation on task resolution
(Table~\ref{tab:main-results}: all five contrasts non-significant) while carrying roughly
half the memory footprint (Figure~\ref{fig:pareto-footprint}), and it does so at no compute
saving and no compute penalty (Figure~\ref{fig:pareto-compute}: a $\approx 6\%$ near-null).
The case for forgetting here is therefore a footprint case, not a resolution case and not a
compute case. Because the resolution proxy is itself coarse, the rest of the chapter
quantifies \emph{why} the value-axis gap is as small as it is --- the task-level
binomial/logit GLMM (Section~\ref{sec:h4}), the per-sequence memory-lift heterogeneity
(Section~\ref{sec:h5}), and the A5 anchor sub-study that probes the forgetting dynamics the
resolved-rate proxy cannot see.
