% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py.

% ── 5.4  H3: Compressive Consolidation (A Negative Result) ───────────────
% PURPOSE: Report, honestly, that CLS Consolidation never fired on SWE-Bench-CL and
%          reframe the non-event as a finding about WHEN consolidation applies.
% MOVE: Honest negative — state the null up front, then mine it for insight.
% BEATS: see ordered bullets below.
% FIGURES/TABLES: none.
% CITATIONS: mcclelland1995 (CLS theory — why consolidation needs redundancy to bite).
% ─────────────────────────────────────────────────────────────────────────

\section{H3: Compressive Consolidation (A Negative Result)}
\label{sec:h3}

\paragraph{The null, stated plainly.}
H3 asked whether the CLS Consolidation policy, which periodically abstracts groups of
related memories into a compact gist, would shift the CL-F1-versus-cost frontier relative
to undifferentiated accumulation. It does not. Across the eight SWE-Bench-CL sequences,
\texttt{cls\_consolidation} resolves $0.449$ of tasks on average against $0.445$ for
\texttt{full\_memory} --- a difference of less than half a percentage point. The
pre-registered Wilcoxon signed-rank test over the $N=8$ sequence-level means (Invariant
\#11) returns $W = 13$, $p = 0.938$, Holm-adjusted $p = 1.000$, with a negligible
rank-biserial effect size $r_{rb} = -0.071$ and a $95\%$ BCa confidence interval on the
paired difference of $[-0.042,\ 0.034]$. The interval straddles zero, the paired median
difference is $-0.0100$, and the Holm-corrected $p$ is at the ceiling. \textbf{H3 is not
supported.} CLS Consolidation is statistically \emph{indistinguishable} from full memory;
it is neither better nor worse. Its marginally higher mean carries no inferential weight
and we make no claim of superiority --- consistent with the chapter's overall result that
no policy separates from full-memory accumulation on this resolved-rate proxy.

Note that CL-F1 here is operationalised as the per-sequence resolved rate, a proxy for
continual-learning performance rather than a direct forgetting curve; a policy could in
principle reshape the memory store in ways this scalar cannot see. The construct-validity
boundary of that proxy is examined separately in the anchor-probe sub-study
(Section~\ref{sec:a5}); the present section concerns only the pre-registered outcome test.

\paragraph{Why consolidation stayed quiet: the gate-3 mechanism.}
The reason for the null is mechanical rather than statistical, and it was identified during
the gate-3 diagnostics. CLS Consolidation is \emph{scheduled} on a fixed cadence --- the
consolidation routine is invoked every $k=5$ tasks (Invariant \#9), not triggered by store
overflow --- so the policy was given regular opportunities to act. What it lacked was
material to act on. The consolidation step first runs DBSCAN over the stored memory
embeddings and only summarises a cluster once at least three memories fall within a
$\geq 0.70$ cosine neighbourhood. On SWE-Bench-CL that density criterion is rarely met:
even within a single repository (retrieval is same-repo only, Invariant \#16), the stored
experiences span distinct issues, files, and error signatures, so the embeddings are too
heterogeneous to form a dense enough cluster. The clustering step therefore seldom yields a
group large and tight enough to compress, and the consolidation branch fires only
occasionally over the matrix.

\paragraph{Integrity: the threshold was not relaxed.}
The obvious way to manufacture a positive result would have been to lower the $\geq 0.70$
cosine threshold (or the minimum cluster size) until DBSCAN began producing clusters. We
did not do this. The threshold is a frozen pre-registered parameter (Invariant \#23), and
tuning a policy's trigger post hoc to force activation would invalidate the comparison and
the pre-registration it rests on. We report the policy as it was specified and let the null
stand. The decision to preserve the frozen trigger over a flattering but engineered
outcome is itself part of the result.

\paragraph{A boundary condition, not a dead end.}
Read as a boundary condition, the non-event is informative. Complementary Learning Systems
theory motivates consolidation precisely as the slow abstraction of \emph{repeated,
overlapping} experience into a stable gist~\cite{mcclelland1995}; redundancy is the
precondition that makes compression worthwhile. A coding-agent workload in which each task
is a distinct bug in a distinct part of the codebase supplies little such redundancy, so
the precondition for compressive consolidation is largely absent here. This does not refute
the mechanism; it locates where it would and would not be expected to bite. Consolidation
should pay off on workloads with genuinely recurring structure --- repeated invocations of
the same API, families of near-duplicate fixes, or tasks that revisit the same modules ---
and SWE-Bench-CL, by construction, is not such a workload.

\paragraph{Convergent evidence from the task-level model.}
The sequence-level null is echoed at the task level. In the binomial/logit GLMM with a
crossed random effect on \texttt{task\_id} (Invariant \#14, $4{,}914$ observations),
\texttt{cls\_consolidation} serves as the reference policy. With CLS as the baseline, none
of the remaining policy coefficients is significant (every $|\text{coef}| < 2\,\mathrm{SE}$:
\texttt{full\_memory} $+0.116 \pm 0.118$, \texttt{no\_memory} $-0.133 \pm 0.117$,
\texttt{random\_prune} $+0.033 \pm 0.117$, \texttt{recency\_prune} $-0.092 \pm 0.117$,
\texttt{type\_aware\_decay} $-0.188 \pm 0.117$). No policy is distinguishable from CLS once
task difficulty and sequence position are accounted for, corroborating the H3 null from an
independent modelling angle. (The GLMM fit is approximate, obtained via a
\texttt{statsmodels} variational procedure rather than the canonical \texttt{lme4}
estimator; we therefore report coefficient signs and the null, and a sensitivity check
confirms the direction.)
