% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py.

% ── 5.7  Manipulation Checks ─────────────────────────────────────────────
% PURPOSE: Prove the storage manipulation actually reaches the prompt before any outcome
%          is interpreted — the experiment must be visible at the point of injection.
% MOVE: Pain → Consequence → Gap (a manipulation you can't see is a confound you can't rule out).
% FIGURES/TABLES: none.
% CITATIONS: none new.
% ─────────────────────────────────────────────────────────────────────────

\section{Manipulation Checks}
\label{sec:manipulation-checks}

The hypothesis tests in Sections~\ref{sec:h1}--\ref{sec:h5} are only interpretable
if three things actually held during execution: the pruning policies genuinely
discarded memory, every condition retrieved through the same scoring path, and the
underlying model and embedder never moved. A manipulation that cannot be seen at the
point of injection is a confound that cannot be ruled out. This section is therefore
the gate the reader must pass through before reading the preceding results causally:
it confirms that the six policies materially differ in what reaches the prompt, that
they differ \emph{only} in the dimension the design intends, and that the recorded
matrix is internally complete. Throughout, the model is held constant at
\texttt{deepseek-v4-flash} (Deviation~D1), so any between-policy difference cannot be
attributed to a moving generator.

\subsection{The pruning manipulation binds}

The first check is that eviction was not vacuous. The per-policy cap is fixed at
$10$ memory items, whereas every SWE-Bench-CL sequence contains far more tasks than
that: the full matrix comprises $4{,}914$ task rows across $8$ sequences, $3$ seeds
and $6$ policies, so each sequence runs many times longer than the cap. The store
therefore overflows the cap well before any sequence completes, and the pruning
policies are forced to make eviction decisions on every subsequent task rather than
sitting idle. The cap binds by construction.

The clearest \emph{quantitative} evidence that the manipulation reaches the prompt is
the realised memory footprint --- the number of memory tokens actually injected per
task, measured at the point of injection rather than inferred from the store. The
policies separate sharply on this axis:

\begin{itemize}
  \item \texttt{full\_memory} injects \textbf{1{,}247} tokens per task --- it carries
        the whole accumulated store;
  \item the four bounded policies inject roughly half as much ---
        \texttt{cls\_consolidation} \textbf{704}, \texttt{recency\_prune}
        \textbf{604}, \texttt{random\_prune} \textbf{601} and
        \texttt{type\_aware\_decay} \textbf{598} tokens per task --- consistent with a
        store held to the cap rather than allowed to grow;
  \item \texttt{no\_memory} injects \textbf{0} tokens, the clean zero-footprint
        boundary condition.
\end{itemize}

A roughly $2\times$ spread between \texttt{full\_memory} and the pruners, and an exact
zero for \texttt{no\_memory}, demonstrate that the policies change \emph{what lands in
the prompt}, not merely what sits in the store. This footprint separation is reported
here purely as a manipulation check --- evidence that the intervention took effect; its
performance consequences (the footprint axis of H1b) are interpreted in
Section~\ref{sec:h1}, not here.

\subsection{Retrieval is identical across policies}

The second check is that the policies differ \emph{only} in eviction. Retrieval
scoring is pure cosine similarity routed through a single shared code path
(\texttt{shared\_retrieve}, Invariant~\#5): no policy adds bonuses, penalties, or
re-ranking. The same embedder (\texttt{nomic-embed-text-v2-moe}, $768$-d,
Deviation~D2) and the same injection ordering --- relevance-sorted with the best item
placed last to mitigate lost-in-the-middle effects (Invariant~\#6) --- are used in
every condition. Consequently the only thing that varies between the six conditions is
\emph{which items survive in the store to be scored}, which is exactly the
manipulation under test. \texttt{no\_memory} is the degenerate case of this same path
with an empty store: it retrieves nothing by construction, giving a clean lower
boundary that no other condition can leak into. The memory-type classifier (D4) is
run identically across all six conditions and its failures are handled the same way in
each, so it cannot differentially bias any single policy.

\subsection{The full matrix audits clean}

The final check is that the recorded matrix is internally consistent before any
statistic is computed on it. A deep audit of the trustworthy run
(\texttt{runs\_144\_seq\_cceb325}, agent frozen at commit \texttt{cceb325}) returned
\textbf{zero flags} across all $4{,}914$ task rows, of which $2{,}175$ ($44.3\%$) were
resolved. The audit confirmed in particular that:

\begin{itemize}
  \item \textbf{Exact task-id sets.} Every policy and seed was evaluated on the
        identical set of tasks --- no condition received an easier or smaller subset,
        which is itself a manipulation check against selection differences across
        arms.
  \item \textbf{Frozen agent.} The coding agent was pinned at commit
        \texttt{cceb325} for the entire run; \texttt{git diff cceb325..HEAD --
        src/agents} is empty, so the agent code is provably constant across all $144$
        units.
  \item \textbf{Tool-mode consistency.} The per-row tool-mode mismatch that affected
        an earlier matrix was gated and blocked before this run, so no row mixes
        tool-calling conventions.
\end{itemize}

For full transparency, two execution facts are disclosed even though neither affects
the science. First, the orchestration code changed during the run, but only for
disk management --- a per-task \texttt{docker rmi} cleanup gated behind the
\texttt{CLEANUP\_TASK\_IMAGES} environment flag (covered by
\texttt{tests/test\_cleanup\_task\_image.py}); the agent, memory, and evaluation paths
were untouched. Second, one cell (\texttt{random\_prune\_matplotlib\_seed2}) had its
\texttt{RUN\_COMPLETED} marker written by hand \emph{after}
\texttt{validate\_run\_complete} had already passed on its data: the marker file was
lost to a process kill, the underlying outputs are complete, and the hand-written
marker only re-records a completion that the validator had independently confirmed.

\subsection{Provenance}

Each unit's manipulation is auditable after the fact: the model
(\texttt{deepseek-v4-flash}), embedder, seed, cap, and policy configuration are
recorded per run, with model provenance fixed through the run directory plus the
frozen \texttt{base.yaml} and \texttt{MODEL\_PROVENANCE.json}. Taken together, these
checks establish the precondition for reading Sections~\ref{sec:h1}--\ref{sec:h5}
causally: the manipulation is real and visible at the prompt, the conditions differ
only in eviction, and the matrix on which every statistic is computed is complete and
flag-free.
