% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py (prose drafted 2026-06-25).

% ── 1.2  Problem Statement ──────────────────────────────────────────────────
% PURPOSE: Name the pain, its consequence, and the unmet research gap.
% MOVE: Pain -> Consequence -> Gap ("therefore there exists a need ...").
% ─────────────────────────────────────────────────────────────────────────

\section{Problem Statement}

The agent designer faces a genuine dilemma with two real failure modes. Keep all memory,
and the context window bloats: token cost and latency climb with every completed task, and
the prompt fills with material that is at best irrelevant and at worst
distracting~\cite{liu2024,du2025}. Prune aggressively, and the agent may discard exactly
the record it needed --- the prior fix to the very subsystem the current issue touches.
Neither extreme is safe, and the harm is not only economic: because agents follow
retrieved experience even when it is wrong, a larger store can lower the resolution rate
rather than raise it~\cite{xiong2025}. ``More memory'' is therefore not a free lever, and
``less memory'' is not a free saving; both directions carry risk.

The consequence is that the field lacks a clean answer to a basic engineering question:
how much does memory actually help, and which forgetting policy --- if any --- best
navigates the trade-off? The deeper obstacle is methodological. In most prior systems the
storage policy is \emph{entangled} with the retrieval policy: a method that ``remembers
better'' usually also retrieves differently, embeds differently, or ranks differently, so
any measured benefit cannot be attributed to forgetting as opposed to search. No benefit
is cleanly attributable, and for coding agents specifically, none has been demonstrated
under a controlled comparison at all.

The closest prior comparison underscores the gap rather than closing it. Alqithami's
six-policy cognitive-memory study~\cite{alqithami2025} evaluates a comparable breadth of
forgetting behaviours, but in a \emph{conversational}, privacy-aware setting and without
holding retrieval constant across policies --- so its conditions differ in both what they
store and how they search, and its domain is not software engineering. It establishes that
forgetting policies are worth comparing; it does not establish what they do for a coding
agent when retrieval is fixed.

Therefore there exists a need for a controlled, \emph{retrieval-fixed} comparison of
forgetting policies for coding agents --- one in which the only thing that varies between
conditions is what the agent keeps versus forgets, so that the effect of the storage
policy can be measured rather than inferred. Supplying that control is the central problem
this thesis addresses.
