% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py (prose drafted 2026-06-25).

% ── 1.4  Contributions ──────────────────────────────────────────────────────
% PURPOSE: Enumerate the thesis contributions and position against prior work.
% MOVE: Pivot ("Rather than invent a memory policy...") + "In contrast".
% Contributions framed as METHOD; finding-clauses left as placeholders.
% ─────────────────────────────────────────────────────────────────────────

\section{Contributions}

Rather than invent yet another memory policy and argue that it wins, this thesis is a
controlled \emph{measurement} of forgetting under one decisive control. Its contributions
are methodological and empirical, and they hold regardless of which way the headline
result points:

\begin{enumerate}
  \item \textbf{Retrieval-held-constant isolation of the storage policy.} The central
        confound control: all six policies retrieve through identical pure-cosine search
        at the same depth, so any between-policy difference is attributable to what is
        \emph{kept} rather than to how it is \emph{searched}. To our knowledge this is the
        first such isolation for coding agents.
  \item \textbf{A No-Memory $\leftrightarrow$ Full-Memory bracketing with a
        pre-registered, effect-size-first analysis.} The two extremes bracket the four
        intermediate policies, and the analysis leads with rank-biserial effect sizes and
        bootstrap confidence intervals rather than dichotomous significance, with the
        primary test (Wilcoxon signed-rank on $N=8$ sequence means, Holm-corrected) and a
        power characterisation fixed in advance.
  \item \textbf{Two-axis Pareto analysis of resolution against both compute and memory
        footprint.} Separating the cost question into a compute axis (total tokens) and a
        storage \emph{footprint} axis (memory carried per task) localises where, if
        anywhere, forgetting pays off (Figure~\ref{fig:pareto-footprint});
        the quantitative frontier is reported in
        Chapter~\ref{ch:results} \texttt{[RESULT: pending runs\_144\_seq\_cceb325]}.
  \item \textbf{A quantification of task interdependence as the precondition for memory to
        matter.} Only a minority of SWE-Bench-CL tasks --- on the order of one-third ---
        structurally depend on a predecessor (a stored record whose gold-patch files
        overlap the current task), which bounds how much any memory policy could help and
        contextualises the experience-following risk~\cite{xiong2025}; the exact figures
        are reported in Chapter~\ref{ch:results}.
  \item \textbf{An instrument-validation sub-study.} A pre-registered A/B health check of
        the agent's own tooling caught a defect in the edit/read instrumentation, drove a
        root-caused fix, and re-validated the corrected instrument before the full matrix
        was run (Section~\ref{sec:instrument}). The trustworthiness procedure is itself a
        reusable methodological contribution.
  \item \textbf{A reproducible artifact.} The full 144-run matrix, the six policy
        implementations, the configuration, and the analysis pipeline are released so the
        measurement can be re-derived end to end (Appendix~\ref{ch:appendix}).
\end{enumerate}

In contrast to the closest prior work, the comparison here moves only one thing.
Alqithami's six-policy study~\cite{alqithami2025} is conversational and does not hold
retrieval fixed, so its conditions differ in both storage and search; coding-oriented
memory systems, in turn, typically vary retrieval granularity or ranking alongside
storage. By freezing retrieval, the model, the dataset, and the agent loop, this thesis
makes the storage policy the only independent variable --- which is what licenses reading
the result as an effect of forgetting itself.
