% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py (prose drafted 2026-06-25).
% SOURCE: AMENDMENTS.md (authoritative D1-D8, A1-A7); docs/amended-gate-2026-06-24.md;
% docs/MORNING-REPORT-2026-06-25.md (disk-capacity / sequence-phased execution).
% INTEGRITY: final model = deepseek-v4-flash (D8, single model all roles). A5 is RESOLVED
% (option c), NOT pending. A6 withdrawn / A7 restored 3 seeds. Hotfix cceb325 is an amendment.

% ── 6.3  Deviations from Pre-Registration ────────────────────────────────
% PURPOSE: Disclose every deviation and amendment as the paper trail against goalpost-moving.
% ─────────────────────────────────────────────────────────────────────────

\section{Deviations from Pre-Registration}
\label{sec:deviations}

The pre-registration is \texttt{THESIS\_FINAL\_v5}; every change from it is recorded here so
the study can be audited against the charge of moving the goalposts after seeing the data.
Each entry passes the same four-part test: it is (a) a mechanical fix to an inconsistency a
prior deviation introduced, or a construct correction the mechanism demands; (b) dated and
decided \emph{before} the affected data was collected; (c) held constant across all
conditions and seeds, so it is a fixed factor that cannot bias a between-policy contrast; and
(d) disclosed. Table~\ref{tab:deviations} lists them; the text foregrounds the load-bearing
ones.

\paragraph{Runtime deviations (D1--D8).}
The most consequential deviation is the model. The pre-registered GPT-5.4 was unavailable
under the project's resource constraints, and the executed model arrived at through a
\emph{disclosed chain} of OpenAI-compatible substitutions: GPT-5.4 (pre-registered) $\to$
Kimi-k2.6 (D1) $\to$ MiniMax M3 (D6) $\to$ a per-role Kimi-plus-DeepSeek split (D7) $\to$,
finally, a \emph{single} model serving all roles --- \texttt{deepseek-v4-flash} on OpenCode,
agent and classifier and reflection and CLS summariser alike (D8). The final single-model
configuration is the cleanest of the chain: it removes the cross-vendor model mix entirely,
so within every cell all six policies share identical models and the model never confounds a
between-policy contrast. Because the model is substituted, absolute resolution rates are
\emph{not} comparable to GPT-5.4 or to public SWE-Bench leaderboards --- only the
between-policy contrasts on the fixed model are claimed. The remaining runtime deviations are
infrastructural and likewise uniform across conditions: the embedder is a local
\texttt{nomic-embed-text-v2-moe} (768-dimensional, D2); cost is measured as a token-count
proxy because the provider is flat-rate (D3); structured outputs use prompt-instructed JSON
with tolerant extraction and Pydantic validation rather than a native JSON-schema mode the
provider does not support (D4/D6b); the host is an x86\_64 cloud fleet rather than the
pre-registered machine (D5); reasoning tokens are stripped before parsing, embedding, and
re-injection (so the memory payload stays reasoning-free), while the cost axis carries a
uniform, between-condition-neutral inflation from the light reasoning model (D6a/D8a); and a
rotating key pool fails closed on quota exhaustion so balance depletion can never be recorded
as a false \texttt{resolved=0} (D6c/D8b). Every runtime deviation is configuration-only and
reversible via environment variables.

\paragraph{Pre-registration amendments (A1--A7).}
Several amendments correct inconsistencies the deviations or the benchmark exposed. The
memory budget cap was lowered from 100 to 10 records (A1) because all eight sequences are
short enough that a cap of 100 never binds --- with it, the four pruning policies would have
been operationally identical to Full Memory, and H2 would be untestable; a cap of 10, below
the shortest sequence, makes pruning genuinely fire. Temperature was raised from 0 to 1 (A2)
because the reasoning endpoint rejects a temperature of 0; it is held constant across all
conditions, and the resulting classifier non-determinism is logged as a failure rate. The
CLS old-memory threshold was lowered to half the cap (A3), a derived consequence of A1, so
that consolidation could in principle select candidates. The H1b efficiency hypothesis was
reframed from token savings to memory footprint and retrieval latency (A4): with retrieval
depth fixed, storage pruning cannot reduce the prompt budget, so footprint is the honest,
mechanism-aligned axis. The anchor-probe scope (A5) was resolved by demoting the full
matrix-scale anchor-probe in favour of a targeted construct-validity sub-study and reporting
CL-F1 as a resolution-rate proxy (Section~\ref{sec:threats}). Finally, a deadline-driven
reduction to a single seed (A6) was \emph{withdrawn} and the full three-seed, 144-run design
restored (A7) once the move to an unlimited provider and a horizontal fleet removed the
compute pressure.

\paragraph{Instrument and execution amendments.}
Two further amendments arose during the rerun and are disclosed here for completeness. First,
the instrument hotfix (commit \texttt{cceb325}): the edit-tool path-normalisation bug
described in Section~\ref{sec:instrument} was fixed, and the pre-registered health gate was
amended on first principles (instrument-attributable failures to zero plus a model-quality
non-regression bound, replacing the absolute total-ratio criterion) and Codex-reviewed before
any re-run data was collected. Second, an execution amendment forced by infrastructure: the
per-task Docker evaluation images accumulate beyond the fleet's disk capacity when all eight
sequences run at once, so the fixed-code matrix is executed under a \emph{sequence-phased}
scheduler --- one sequence's images are pulled, its cells run, and the images pruned before
the next phase --- at the frozen experiment commit \texttt{cceb325}. This changes only the
scheduling of identical work, not what is computed, and is therefore non-confounding.

\begin{table}[htbp]
  \centering
  \caption{Deviations (D) and amendments (A) from the \texttt{THESIS\_FINAL\_v5}
  pre-registration. Every entry is held constant across all conditions and seeds.}
  \label{tab:deviations}
  \begin{tabular}{lll}
    \toprule
    ID & Change & Status / rationale \\
    \midrule
    D1--D8 & model $\to$ \texttt{deepseek-v4-flash} (single, all roles) & fixed factor; absolute rates not comparable \\
    D2 & embedder $\to$ local \texttt{nomic-embed-text-v2-moe} (768-d) & uniform across conditions \\
    D3 & cost metric $\to$ token-count proxy & flat-rate provider \\
    D4/D6b & structured output $\to$ JSON + Pydantic & provider lacks JSON-schema mode \\
    D5 & host $\to$ x86\_64 cloud fleet & uniform; exclusions disclosed \\
    D6a/D8a & reasoning tokens stripped; cost-axis inflation & uniform, between-condition-neutral \\
    A1 & record cap $100 \to 10$ & makes pruning bind; H2 testable \\
    A2 & temperature $0 \to 1$ & endpoint rejects 0; classifier rate logged \\
    A3 & CLS old-memory threshold $\to$ cap/2 & derived from A1 \\
    A4 & H1b $\to$ footprint + latency & mechanism-aligned efficiency axis \\
    A5 & anchor-probe $\to$ construct-validity sub-study & CL-F1 reported as resolve-rate proxy \\
    A6 $\to$ A7 & seeds $1 \to 3$ (A6 withdrawn) & full 144-run design restored \\
    \texttt{cceb325} & edit-tool hotfix + amended gate & pre-registered, Codex-reviewed (\S\ref{sec:instrument}) \\
    exec & sequence-phased scheduling & disk capacity; reschedules identical work \\
    \bottomrule
  \end{tabular}
\end{table}
