% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py (prose drafted 2026-06-25).
% SOURCE: docs/threats-to-validity.md (setup) + docs/a5-anchor-probe-decision.md /
% AMENDMENTS.md A5 (RESOLUTION). INTEGRITY: carry A5 THROUGH -- the instrument is NOT
% measurement-blind; the 1.000 is regime-driven. Model = deepseek-v4-flash. No TOST claim.
% Cite only bib keys (drop the scaffold's [VERIFY+ADD] refs not in references.bib).

% ── 6.4  Threats to Validity ─────────────────────────────────────────────
% PURPOSE: Reframe the study's limits as conscious trade-offs across four validity types.
% ─────────────────────────────────────────────────────────────────────────

\section{Threats to Validity}
\label{sec:threats}

\paragraph{Construct and measurement validity.}
The continual-learning metric's stability dimension is the most delicate construct in this
study. The pre-registered stability instrument --- anchor-probe re-evaluation of earlier
tasks against later memory states --- saturated at $1.000$ on the small gate-3 anchor
sample, which raised a real worry: a saturated metric could mean genuine absence of
forgetting \emph{or} an instrument too insensitive to register it. We resolved this rather
than leaving it open (amendment A5). A targeted discrimination sub-study, run on cells
selected so that Full Memory and an aggressive pruner actually retrieve different records,
shows the instrument is \emph{not} measurement-blind: on cells where Full retained and
retrieved a task's own solution memory while the pruner had evicted it, Full re-solved the
task where the pruner could not (a clean existence proof of one cell at three-of-three versus
zero-of-three). The saturation is therefore \emph{regime-driven}, not a broken metric: in the
natural SWE-Bench-CL regime most tasks are solvable without memory, or the relevant memory is
not cosine-retrieved, so stability is rarely exercised. Two limits bound this sub-study and
are stated narrowly: its signal is \emph{backward retention only} --- the forward, cross-task
interdependence cells were at the floor in both conditions, so it gives no support for
``earlier-task memory helps later tasks'' --- and it carries no p-value, because the cells
were selected to maximise discrimination. Consequently CL-F1 is reported as a
resolution-rate proxy and the efficiency claim is carried by the directly-measured footprint
axis, which is immune to this sensitivity. Relatedly, CLS Consolidation never clustered on
the benchmark's heterogeneous same-repository memories, so H3 is reported as not-applicable
rather than supported; we declined to relax the frozen similarity threshold, which would have
manufactured incoherent summaries. Finally, because the executed model emits reasoning tokens
counted in completion length, the token-cost axis is inflated relative to a non-reasoning
model; the inflation is uniform across conditions, so between-policy cost contrasts hold
while absolute token costs do not transfer across studies.

\paragraph{Internal validity.}
Retrieval is held identical --- pure cosine, fixed depth --- across all six conditions, so
the contrast is attributable to the storage and forgetting policy rather than to search. The
model, embedder, temperature, and memory budget are fixed factors held constant across
conditions and seeds. The principal residual threat is benchmark contamination: SWE-Bench
tasks are public and the model's training cutoff is undocumented. Because any contamination
is constant across conditions, it does not bias the between-policy contrasts; it does render
absolute solve rates uninterpretable as capability measures, which we disclaim accordingly.
Per-run provenance --- model alias, configuration hash, and dataset revision --- is logged so
that any drift under the rotating-key router would be detectable. A more concrete internal
threat was the agent's own tooling: a trajectory audit found the \texttt{read\_file} and
\texttt{edit\_file} instruments defective, which could have turned an instrument artefact into
an apparent between-policy effect. This is a \emph{mitigated} threat rather than a live one
--- the defect was root-caused, fixed (commit \texttt{cceb325}), and the corrected instrument
was re-validated by a pre-registered, externally reviewed health gate that must pass before
the matrix is trusted (Section~\ref{sec:instrument}).

\paragraph{External validity.}
The results rest on a single base model (\texttt{deepseek-v4-flash}, the final link of the
disclosed deviation chain), a single benchmark family (SWE-Bench-CL), same-repository
retrieval only, and a fixed retrieval depth. This is a conscious scope boundary, defended by
scope and measurement cleanliness rather than by budget --- the move to an unlimited provider
and a horizontal fleet removed compute as the binding constraint. We make no cross-model or
cross-benchmark generalisation claim. A deeper boundary is intrinsic to the task
distribution: only a minority of tasks structurally depend on a predecessor (a
benchmark-structural property; the exact fractions are reported in Chapter~\ref{ch:results}),
so any conclusion about memory's \emph{benefit} applies to that interdependent subset rather
than to sequential coding in general, and the
experience-following risk~\cite{xiong2025} means an indiscriminate store can hurt the
remainder.

\paragraph{Statistical-conclusion validity.}
The primary test is the Wilcoxon signed-rank test on $N=8$ sequence-level means with Holm
correction over the five pre-registered contrasts, accompanied by rank-biserial effect sizes
and BCa bootstrap intervals. With $N=8$ the design is honestly underpowered for small
effects, which we address by leading with estimation --- effect sizes and intervals, plus the
design's detectable-effect profile --- rather than dichotomous significance, and by not
switching tests to manufacture a significant verdict (Invariant~\#11). Run-level variance
from the agent's stochasticity is estimated from the three seeds and reported separately from
task-level bootstrap intervals; the two are distinct variance sources and are never pooled.
