% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py (prose drafted 2026-06-25).

% ── 2.1  Memory and Continual Learning in LLM Agents ─────────────────────────
% PURPOSE: Frame the thesis within continual/sequential learning and demarcate external
%          non-parametric memory (our subject) from weight-level CL.
% ─────────────────────────────────────────────────────────────────────────────

\section{Memory and Continual Learning in LLM Agents}

An LLM coding agent can resolve a real repository issue, but resolving issues one at a time
is not the same as \emph{learning across} a stream of them. The capability --- plan, edit,
test, iterate --- is mature; the policy that governs what an agent should carry from one
task to the next is not. This is the setting of continual, or sequential,
learning~\cite{joshi2025}: an agent works a stream of tasks and ought to bring useful
experience forward without re-training, yet in practice the decision of \emph{what} to
carry is left unprincipled.

The benchmark used here, SWE-Bench-CL, is a \emph{task-incremental} continual-learning
setting: the tasks within each sequence arrive in a fixed order (the eight official
sequences, never reordered), and a later task may depend on an earlier one in the same
repository. The outcome the thesis cares about, CL-F1, is designed to balance the two
pulls of continual learning --- \emph{plasticity}, the ability to solve the new task, and
\emph{stability}, the retention of competence on earlier ones; its definition is given
with the other metrics in Table~\ref{tab:metrics}.

There are two distinct places where ``learning across tasks'' can happen. One is
\emph{parametric}: continual learning in the model's weights, where the classic risk is
gradient interference and catastrophic forgetting. The other is \emph{external} and
non-parametric: a frozen model paired with a retrieval-backed store of past experience.
This thesis studies the latter exclusively. The model is held frozen and the only thing
that changes across the six conditions is the storage and forgetting decision applied to
an external \texttt{MemoryStore}. Weight-level continual learning and catastrophic
forgetting in the parameters are explicitly out of scope --- we never touch the model
weights --- which keeps the object of study cleanly the memory policy rather than the
learning dynamics of the network.
