% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py (prose drafted 2026-06-25).

% ── 2.2  The Plasticity-Stability Tradeoff and Forgetting ────────────────────
% PURPOSE: Establish the stability/plasticity tension and the counter-intuitive claim
%          that forgetting can HELP, motivating proactive memory policies.
% ─────────────────────────────────────────────────────────────────────────────

\section{The Plasticity--Stability Tradeoff and Forgetting}

Contrary to the intuition that more memory is always better, an agent that retains
everything can be worse off than one that forgets. The reason is the classic
plasticity--stability tension: retaining prior competence pulls against adapting to what is
new. An unbounded store maximises stability in principle, but in doing so it risks drowning
the one relevant record among a growing mass of irrelevant ones.

In parametric learning this tension surfaces as catastrophic forgetting --- new gradients
overwrite old competence --- but that framing is at the level of weights. For an external,
retrieval-backed memory the failure mode inverts: the danger is not erasure but
\emph{accumulation}. Stale or off-task records do not simply sit harmlessly in the store;
they act as distractors that crowd out the relevant item and bias what retrieval surfaces.
The cost of remembering is paid not only in tokens but in the quality of what is retrieved.

This is where forgetting becomes a design lever rather than a deficiency. Empirically,
agents exhibit \emph{experience-following}: a retrieved precedent is reused faithfully,
which is a virtue when the precedent is apt and a liability when it is not. A bad stored
experience, once retrieved, is reused and its error compounds across subsequent
tasks~\cite{xiong2025}. Under experience-following, pruning a misleading record is not lost
information --- it is removed interference. Proactive forgetting can therefore raise
performance, not merely lower cost.

This reframes the whole question the thesis asks. If forgetting can help, then the design
problem is not how to remember more but \emph{when} and \emph{how much} to forget --- and
the answer is bounded by how often memory is even relevant. Most coding tasks in the
benchmark do not share files with any predecessor (Figure~\ref{fig:interdependence}), so
for the majority of tasks there is nothing useful to carry forward and accumulation buys
only risk. Not every task needs memory; the policies in this thesis are different bets on
which records are worth keeping for the minority that do.
