% Generated from the matching Typst scaffold by
% scripts/migrate_typst_report_to_latex.py (prose drafted 2026-06-25).

% ── 2.4  Retrieval-Augmented Memory and Context Rot ──────────────────────────
% PURPOSE: Explain dense retrieval over the store and why retrieval is HELD CONSTANT
%          across all six policies to isolate the storage/forgetting decision.
% ─────────────────────────────────────────────────────────────────────────────

\section{Retrieval-Augmented Memory and Context Rot}

A memory store is only as useful as what retrieval pulls from it. In this work retrieval is
dense: each stored record is embedded and scored against the current query by cosine
similarity, and the top $k=5$ same-repository records are injected into the agent's prompt.
The embedded payload is deliberately compact --- the issue text, the final error, and the
final diff, kept under the embedder's token cap --- and embeddings come from a local
\texttt{nomic-embed-text-v2-moe} model (see the deviations in Section~\ref{sec:deviations}).

The pain is that retrieval is not neutral about \emph{where} in the prompt a record lands.
Models use the middle of a long context least --- the lost-in-the-middle
effect~\cite{liu2024} --- and the presence of plausible-but-irrelevant retrieved passages
can itself change the answer~\cite{cuconasu2024}. The agent mitigates the positional part
by ordering the injected records so the best-scoring one comes \emph{last}, nearest the
generation point.

The consequence runs deeper than position. Even with retrieval held perfect, simply
lengthening the context degrades performance: irrelevant tokens accumulated in the window
lower answer quality regardless of whether the relevant item is present~\cite{du2025}. For
an accumulating memory store this is the quiet tax on remembering --- every kept record is
a candidate distractor, so the storage policy and the retrieval outcome are coupled.

This coupling is exactly the gap the thesis's design closes. Because retrieval is itself a
powerful confound --- a policy could appear to ``remember better'' merely by searching
differently --- retrieval is \emph{held constant across all six policies}: pure cosine
similarity, the same depth, the same context budget, same-repository only. With search
frozen, any difference between conditions is attributable to what each policy stores and
forgets, not to how it retrieves. Figure~\ref{fig:architecture} shows where this control
sits in the pipeline: the storage policy is the only swappable component.
